---
title: "chem"
author: "Guangyu Ding"
date: "2024-10-23"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Consider the survey_d_chem file. 

```{r}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)

chem_data <- read_csv("survey_d_chem.csv")

## Filter data for California, because California is one of the major strawberry 
## producing regions in the United States. Observing the data here is more 
## convincing.
chem_data_ca <- chem_data %>%
  filter(State == "CALIFORNIA")

# Clean the data by removing non-numeric values and converting to numeric
chem_data_ca <- chem_data_ca %>%
  filter(!Value %in% c("(D)", "(NA)")) %>%
  mutate(Value = as.numeric(Value))

# Summarize total applications per year for each chemical category
chem_summary <- chem_data_ca %>%
  group_by(Year, col2) %>%
  summarise(Total_Value = sum(Value, na.rm = TRUE)) %>%
  spread(key = col2, value = Total_Value)

# Print the summary
print(chem_summary)

## Visualize the data by making line chart 
chem_data <- chem_data %>%
  filter(!Value %in% c("(D)", "(NA)")) %>%
  mutate(Value = as.numeric(Value))

# Summarize total applications per year across all chemical types
chem_summary <- chem_data %>%
  group_by(Year) %>%
  summarise(Total_Value = sum(Value, na.rm = TRUE))

ggplot(chem_summary, aes(x = Year, y = Total_Value)) +
  geom_line(color = "darkgreen", size = 1) +
  geom_point(color = "darkgreen", size = 2) +
  labs(title = "Total Chemical Application Trends",
       x = "Year",
       y = "Total Applications (in pounds)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Now separate the values by the categories listed in col2 column. Through 
## this we can see which chemicals are being used more every year. 

```{r}
# Summarize total applications per year for each chemical category (col2)
chem_summary1 <- chem_data %>%
  group_by(Year, col2) %>%
  summarise(Total_Value = sum(Value, na.rm = TRUE))

# Spread the data to have categories as columns (if needed)
chem_summary_wide <- chem_summary1 %>%
  spread(key = col2, value = Total_Value)

# Print the summary data for each chemical category
print(chem_summary1)
## Obviously, Insecticides are the most commonly used chemical substances for 
## each year.

## We can see the trend of usage of each types of chemical substances by making 
## a line chart to show all types of chemical substances used from 2018 to 2023.
chem_data_ca <- chem_data_ca %>%
  filter(!Value %in% c("(D)", "(NA)")) %>%
  mutate(Value = as.numeric(Value))

chem_summary <- chem_data_ca %>%
  group_by(Year, col2) %>%
  summarise(Total_Value = sum(Value, na.rm = TRUE)) %>%
  spread(key = col2, value = Total_Value)

chem_long <- chem_summary %>%
  gather(key = "Chemical_Type", value = "Total_Value", -Year)

ggplot(chem_long, aes(x = Year, y = Total_Value, color = Chemical_Type, group = Chemical_Type)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Chemical Application Trends in California",
       x = "Year",
       y = "Total Applications (in pounds)",
       color = "Chemical Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

## Based on the plot, the usage of insecticide decreases from 2018 to 2023.
## The fungicide had an increase in 2019, but decreased since then.
## The herbicide has an overall trend of decreasing, but experienced a rebound in 2021.
## All other chemical substances has a similar trend with fungicide, but decreased more from 2019 to 2021, less from 2021 to 2023.


## We can make more specific observation on each types of chemical substances. i.e. Fungicide
# Filter data for California and fungicides
fungicide_data_ca <- chem_data %>%
  filter(State == "CALIFORNIA", col2 == "FUNGICIDE")

# Clean the data by removing non-numeric values and converting to numeric
fungicide_data_ca <- fungicide_data_ca %>%
  filter(!Value %in% c("(D)", "(NA)")) %>%
  mutate(Value = as.numeric(Value))

# Summarize total applications per year for fungicides
fungicide_summary <- fungicide_data_ca %>%
  group_by(Year) %>%
  summarise(Total_Value = sum(Value, na.rm = TRUE))

# Plot the trends using ggplot2
ggplot(fungicide_summary, aes(x = Year, y = Total_Value)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "blue", size = 2) +
  labs(title = "Fungicide Application Patterns in California",
       x = "Year",
       y = "Total Applications (in pounds)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Now consider the suevey_d_fert file.By observing this file, we can get the usage of the fertilizer

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)

fert_data <- read_csv("survey_d_fert.csv")

fert_data <- fert_data %>%
  filter(!Value %in% c("(D)", "(NA)")) %>%
  mutate(Value = as.numeric(Value))

# Summarize total applications per year for each fertilizer type
fert_summary <- fert_data %>%
  group_by(Year, chem_name) %>%
  summarise(Total_Value = sum(Value, na.rm = TRUE))

# Convert data to long format for easier plotting
fert_long <- fert_summary %>%
  spread(key = chem_name, value = Total_Value)

# Plot the trends
ggplot(fert_summary, aes(x = Year, y = Total_Value, color = chem_name, group = chem_name)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Fertilizer Application Trends",
       x = "Year",
       y = "Total Applications (in pounds)",
       color = "Fertilizer Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## By observation, from March 2018 to June 2020, nitrogen was the one being used 
## most. For the rest of time period from 2018 to 2023, Potash was at the first place.

## Now visualize the usage of fertilizer in California. Compare the overall data 
## in the United States with the data in California.

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)

fert_data_ca <- read_csv("fert_data_ca.csv")

summary(fert_data_ca)
str(fert_data_ca)
head(fert_data_ca)
fert_data_ca <- fert_data_ca %>%
  filter(!Value %in% c("(D)", "(NA)")) %>%
  mutate(Value = as.numeric(Value))
fert_summary <- fert_data_ca %>%
  group_by(Year, chem_name) %>%
  summarise(Total_Value = sum(Value, na.rm = TRUE))

print(fert_summary)

# Visualize trends in fertilizer applications over the years
ggplot(fert_summary, aes(x = Year, y = Total_Value, color = chem_name, group = chem_name)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Fertilizer Application Trends in California",
       x = "Year",
       y = "Total Applications (in pounds)",
       color = "Fertilizer Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## We find that the usage of Nitrogen is ahead in California. Maybe other places 
## in United State can change from using potash to nitrogen. 

## Now consider the files fung, fung_ca_only and fung_fl_only to observe which chem is used most.

```{r}
library(dplyr)
library(readr)
library(ggplot2)

fung_fl <- read_csv("fung_fl_only.csv")
fung_data <- read_csv("fung.csv")
fung_ca <- read_csv("fung_ca_only.csv")

summary(fung_fl)
summary(fung_data)
summary(fung_ca)

# Combine the data into one DataFrame for analysis
combined_fung <- bind_rows(fung_fl, fung_data, fung_ca)

# Summarize or clean the data if necessary (e.g., handle missing values)
combined_fung <- combined_fung %>%
  filter(!is.na(chem_index), !is.na(chem_name))

# Create a bar plot to visualize chem_name versus chem_index
ggplot(combined_fung, aes(x = chem_name, y = chem_index, fill = chem_name)) +
  geom_bar(stat = "identity") +
  labs(title = "Chem Index Values by Chem Name",
       x = "Chem Name",
       y = "Chem Index") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3")
```

## The plot produced seems not that useful. Here I think what we need to do is 
## to sort the data in the three fung tables in descending order to obtain the 
## usage quantities of each chemical substance nationwide, in California, and 
## in Florida.

## To prove the series of results above, we can directly observe the relevant 
## data of the harvest quantity.The files sur_CA_19_23 and sur_fl_19_23 are used.
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)

# Load the data from the CSV files
data_ca <- read_csv("sur_CA_19_23.csv")
data_fl <- read_csv("sur_fl_19_23.csv")

# Extract the years for plotting
years <- c("value_19", "value_20", "value_21", "value_22", "value_23")

# Convert the selected columns to numeric after removing commas
convert_to_numeric <- function(x) as.numeric(gsub(",", "", x))

# Visualize PRICE RECEIVED trends for California and Florida
price_ca <- data_ca %>% filter(product_price == "PRICE RECEIVED") %>% select(all_of(years)) %>% unlist() %>% as.numeric()
price_fl <- data_fl %>% filter(product_price == "PRICE RECEIVED") %>% select(all_of(years)) %>% unlist() %>% as.numeric()

# Plot PRICE RECEIVED trends
plot_data <- data.frame(
  Year = rep(years, 2),
  Price = c(price_ca, price_fl),
  State = rep(c("California", "Florida"), each = length(years))
)

ggplot(plot_data, aes(x = Year, y = Price, group = State, color = State)) +
  geom_line() + geom_point() +
  ggtitle("PRICE RECEIVED Trend (2019-2023)") +
  xlab("Year") + ylab("Price ($/CWT)") +
  theme_minimal()

# Visualize HARVESTED acres trends for California and Florida
harvested_ca <- data_ca %>% filter(product_price == "HARVESTED") %>% select(all_of(years)) %>% unlist() %>% convert_to_numeric()
harvested_fl <- data_fl %>% filter(product_price == "HARVESTED") %>% select(all_of(years)) %>% unlist() %>% convert_to_numeric()

# Plot HARVESTED acres trends
harvested_data <- data.frame(
  Year = rep(years, 2),
  Acres = c(harvested_ca, harvested_fl),
  State = rep(c("California", "Florida"), each = length(years))
)

ggplot(harvested_data, aes(x = Year, y = Acres, group = State, color = State)) +
  geom_line() + geom_point() +
  ggtitle("HARVESTED Acres Trend (2019-2023)") +
  xlab("Year") + ylab("Acres") +
  theme_minimal()

# Calculate average HARVESTED acres for each state
avg_harvested_ca <- mean(harvested_ca)
avg_harvested_fl <- mean(harvested_fl)

# Print the averages
avg_harvested_ca
avg_harvested_fl
```
## Maybe one conclustion we can get is for strawberriews cultivation, we may use nitrogen as the fertilizer.

## Since the assignment is related to the chemical usage of cultivating strawberries, several data related to 
## the price of the strawberries may not be consider in my report.  